---
title: "datafest analysis"
output: html_document
date: "2026-03-28"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(scipen=999)
setwd("C:/Users/avril/Downloads/2026-ASA-DataFest-Data-Files/2026-ASA-DataFest-Data-Files")
library(tidyverse)
encounters=read.csv("encounters.csv")
diagnosis=read.csv("diagnosis.csv")
patients=read.csv("patients.csv")
social_determinants=read.csv("social_determinants.csv")
departments=read.csv("departments.csv")
tigercensuscodes=read.csv("tigercensuscodes.csv")
```

Join datasets to create a big one with `encounters`, `diagnosis` and `patients`. Also creates the variable `EverDiagnosed`, which is 1 if a patient was, throughout their journey, diagnosed with a mental illness or disorder.
```{r}
#unique primary diagnosis key
diagnosis_unique = diagnosis %>%
  distinct(DiagnosisKey, .keep_all = TRUE)

encounter_diagnosis = left_join(
  encounters,
  diagnosis_unique,
  by = c("PrimaryDiagnosisKey" = "DiagnosisKey")
)
encounter_diagnosis_patients=left_join(encounter_diagnosis,patients,by=c("PatientDurableKey"="DurableKey"))

encounter_diagnosis_patients <- encounter_diagnosis_patients %>% 
  mutate(MentalDisorder = ifelse(str_starts(DiagnosisValue, "F") & !str_starts(DiagnosisValue, "F7") | str_starts(DiagnosisValue, "R45"),1,0))

encounter_diagnosis_patients <- encounter_diagnosis_patients %>%
  group_by(PatientDurableKey) %>%
  mutate(
    EverDiagnosed = if_else(any(MentalDisorder == 1, na.rm = TRUE), 1, 0)
  ) %>%
  ungroup()

encounter_diagnosis_patients <- encounter_diagnosis_patients %>%
  group_by(PatientDurableKey) %>%
  summarize(
    # Returns TRUE if any encounter for this patient was a 1
    EverDiagnosed = any(MentalDisorder == 1, na.rm = TRUE)
  ) %>%
  mutate(EverDiagnosed = as.numeric(EverDiagnosed))
```
-------------------------------------------------------------------------------
# STATISTICAL ANALYSIS: CHI-SQUARE HYPOTHESIS TESTING
MyChart activity sample plots and Chi-Square tests of independence.
```{r}
#FOR MENTALLY ILL PATIENTS
mental=filter(encounter_diagnosis_patients,EverDiagnosed==1)
analysis_df1 <- mental %>%
  mutate(
    Is_FollowUp = if_else(VisitType == "FOLLOW UP", 1, 0),
        Simple_Status = if_else(MyChartStatus == "Activated", "Active", "Not Active")
  ) %>%
  filter(!is.na(Simple_Status), !is.na(VisitType))

# 1. Create a contingency table
contingency_table1 <- table(analysis_df1$Simple_Status, analysis_df1$Is_FollowUp)
print(contingency_table1)

# 2. Run the Chi-Squared Test
chi_test1 <- chisq.test(contingency_table1)
print(chi_test1)

followup_comparison1 <- analysis_df1 %>%
  group_by(Simple_Status) %>%
  summarize(
    Total_Encounters = n(),
    FollowUp_Count = sum(Is_FollowUp),
    FollowUp_Rate = (FollowUp_Count / Total_Encounters) * 100
  )

print(followup_comparison1)

ggplot(followup_comparison1, aes(x = Simple_Status, y = FollowUp_Rate, fill = Simple_Status)) +
  geom_col(width = 0.6) +
  geom_text(aes(label = paste0(round(FollowUp_Rate, 1), "%")), vjust = -0.5) +
  labs(
    y = "Follow-up Rate (%)",
    x = "MyChart Status"
  ) +
  theme_minimal() +
  guides(fill = "none")


#FOR NON MENTALLY ILL PATIENTS
nonmental=filter(encounter_diagnosis_patients,EverDiagnosed==0)
analysis_df2 <- nonmental %>%
  mutate(
    Is_FollowUp = if_else(VisitType == "FOLLOW UP", 1, 0),
    
    # Simplify MyChart status into two main groups
    Simple_Status = if_else(MyChartStatus == "Activated", "Active", "Not Active")
  ) %>%
  # Remove NAs to keep the statistical test clean
  filter(!is.na(Simple_Status), !is.na(VisitType))

# 1. Create a contingency table
contingency_table2 <- table(analysis_df2$Simple_Status, analysis_df2$Is_FollowUp)
print(contingency_table2)

# 2. Run the Chi-Squared Test
chi_test2 <- chisq.test(contingency_table2)
print(chi_test2)

followup_comparison2 <- analysis_df2 %>%
  group_by(Simple_Status) %>%
  summarize(
    Total_Encounters = n(),
    FollowUp_Count = sum(Is_FollowUp),
    FollowUp_Rate = (FollowUp_Count / Total_Encounters) * 100
  )
```

Measures if follow up visits and mychart activation status are related. Strong relationship found with chi squared hypothesis testing.

```{r}
my_combined <- bind_rows(
  followup_comparison1 %>% mutate(Group = "Mental Illness"),
  followup_comparison2 %>% mutate(Group = "No Mental Illness")
)

ggplot(my_combined, aes(x = Simple_Status, y = FollowUp_Rate, fill = Group)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.7) + 
  geom_text(aes(label = sprintf("%.1f%%", FollowUp_Rate)), 
            position = position_dodge(width = 0.8), 
            vjust = -0.5, 
            fontface = "bold", 
            size = 4) +
  scale_fill_manual(values = c("Mental Illness" = "cornflowerblue", "No Mental Illness" = "darkred")) +
  labs(
    title = "Follow-up Rates by MyChart Status & Mental Health Diagnosis",    y = "Follow-up Rate (%)",
    x = "MyChart Status",
    fill = "Patient Group"
  ) +
  ylim(0, 25) + 
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 16, hjust = 0.5),
    axis.title = element_text(face = "bold", size = 12),
    axis.text = element_text(face = "bold", color = "black", size = 11),
    legend.position = "top",
    legend.title = element_text(face = "bold"),
    panel.grid.major.x = element_blank() # Clean up vertical lines
  )
```

Chi-square test for MyChart inactivation by reported race, including odds ratio computations.

```{r}
# 1. Get the most recent status for each unique patient
patient_mychart_status <- encounter_diagnosis_patients %>%
  group_by(PatientDurableKey) %>%
  slice_max(Date, n = 1, with_ties = FALSE) %>% 
  ungroup()

# 2. Calculate the Inactivation Rate by Race
# 1. Summarize Non-Mental Health Group
inactivation_stats1 <- nonmental %>%
  group_by(FirstRace) %>%
  summarize(
    Total_Patients = n(),
    Inactivated_Count = sum(MyChartStatus != "Activated", na.rm = TRUE),
    Inactivation_Rate = (Inactivated_Count / Total_Patients) * 100,
    Illness = "No Mental Illness"
  )

# 2. Summarize Mental Health Group
inactivation_stats2 <- mental %>%
  group_by(FirstRace) %>%
  summarize(
    Total_Patients = n(),
    Inactivated_Count = sum(MyChartStatus != "Activated", na.rm = TRUE),
    Inactivation_Rate = (Inactivated_Count / Total_Patients) * 100,
    Illness = "Mental Illness"
  )

# 3. Combine into one master dataset
inactivation_stats <- bind_rows(inactivation_stats1, inactivation_stats2) %>%
  arrange(Illness, desc(Inactivation_Rate))
inactivation_stats=inactivation_stats%>%slice(-c(11,20,7,10,17,18))

# 4. Prepare the data for the Chi-square test
chisq_prep <- inactivation_stats %>%
  group_by(FirstRace) %>%
  summarize(
    Inactivated = sum(Inactivated_Count),
    Activated = sum(Total_Patients) - sum(Inactivated_Count)
  )

# 5. Create a Contingency Table
contingency_table <- as.matrix(chisq_prep[, c("Inactivated", "Activated")])
rownames(contingency_table) <- chisq_prep$FirstRace

# 6. Perform the Chi-Square Test
chisq_result <- chisq.test(contingency_table)

# 7. View Results
print(chisq_result)

# 8. Odds Ratios (Minority vs White)
binary_data <- chisq_prep %>%
  mutate(Group = if_else(FirstRace == "White or Caucasian", "White or Caucasian", "Minority")) %>%
  group_by(Group) %>%
  summarize(
    Inactivated = sum(Inactivated),
    Activated = sum(Activated)
  )

binary_table <- as.matrix(binary_data[, c("Inactivated", "Activated")])
rownames(binary_table) <- binary_data$Group

or_test <- (binary_table["Minority", "Inactivated"] / binary_table["Minority", "Activated"]) /
           (binary_table["White or Caucasian", "Inactivated"] / binary_table["White or Caucasian", "Activated"])

cat("Odds Ratio (Minority vs White Inactivity):", or_test, "\n")
```

Creates a bar graph showing the disparities between myChart activation and race.

```{r}
ggplot(inactivation_stats, aes(x = reorder(FirstRace, Inactivation_Rate), y = Inactivation_Rate,group=Illness,fill=Illness)) + scale_fill_manual(values=c("Mental Illness"="cornflowerblue","No Mental Illness"="darkred")) +
  geom_col(position="dodge") +
  coord_flip() +
  labs( title = "MyChart Inactivation Percentages By Reported Race",
        x = "Reported Race",
        y = "Prevalence (% of all patients)") +
  theme_minimal() + 
  theme(
    axis.title = element_text(face = "bold", size = 12),
    plot.title=element_text(face="bold",size=15,hjust=0.5),
    axis.text = element_text(face = "bold", color = "black", size = 12),legend.position = "top"
  ) + ylim(0,45)
```

Creates a bar graph showing the disparities between myChart activation and race.

```{r}
  ggplot(inactivation_stats, aes(x = reorder(FirstRace, Inactivation_Rate), y = Inactivation_Rate,group=Illness,fill=Illness)) + scale_fill_manual(values=c("Mental Illness"="cornflowerblue","No Mental Illness"="darkred")) +
  geom_col(position="dodge") +
  coord_flip() +
  labs( title = "MyChart Inactivation Percentages By Reported Race",
        x = "Reported Race",
        y = "Prevalence (% of all patients)") +
  theme_minimal() + 
  theme(
    axis.title = element_text(face = "bold", size = 12),
    plot.title=element_text(face="bold",size=15,hjust=0.5),
    axis.text = element_text(face = "bold", color = "black", size = 12),legend.position = "top"
  ) + ylim(0,45)
```

Creates a plot showing top diagnosed illnesses and myChart inactivity. Found strong relationship between mental illness (6/10 top cases) and inactivity.
```{r}
disease_inactivation_analysis <- encounter_diagnosis_patients %>%
  filter(!is.na(DiagnosisName)) %>%
  group_by(DiagnosisName) %>%
  summarize(
    Total_Patients = n_distinct(PatientDurableKey),
    Inactivated_Patients = n_distinct(PatientDurableKey[activated_flag==0]),
    Inactivation_Rate = (Inactivated_Patients / Total_Patients) * 100
  ) %>%
  filter(Total_Patients > 500) %>% 
  arrange(desc(Inactivation_Rate)) %>%
  slice_head(n = 10) 
disease_inactivation_analysis <- disease_inactivation_analysis %>%
  mutate(DiagnosisName=c("Disruptive mood dysregulation disorder 
","Major depressive disorder with psychotic symptoms
","Examination for participation in sport
","Suicidal ideations
","Respiratory tuberculosis screening
","Major depressive disorder without psychotic features
","Short stature (children)
","Encounter for pre-employment examination
","Altered mental status, unspecified
","Major depressive disorder (single episode)
"))
plot <- ggplot(disease_inactivation_analysis, aes(x = reorder(DiagnosisName, Inactivation_Rate), y = Inactivation_Rate)) +
  geom_col(fill = "#d95f02", width = 0.7) + 
  coord_flip() +
    geom_text(aes(label = paste0(round(Inactivation_Rate, 1), "%")), 
            hjust = -0.2, 
            size = 3) +
    scale_x_discrete(labels = function(x) str_wrap(x, width = 25)) + 
    scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Top 10 Diagnoses by MyChart Inactivation",
    x = NULL, # Remove "Diagnosis" label to save space
    y = "Inactivation Rate (%)"
  ) +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 8, lineheight = 0.8), # Small text with tight line spacing
    axis.text.x = element_text(size = 8),
    plot.title = element_text(size = 10, face = "bold", hjust = 0.5),
    panel.grid.major.y = element_blank(), # Remove horizontal lines for a cleaner look
    plot.margin = margin(10, 20, 10, 10) # Add extra space on the right for the % labels
  )
ggsave("readable_inactivation_plot.png", plot, width = 6, height = 8, dpi = 300)
```

---------------------------------------------------------------------------------------------------
# EXPLORATORY & SUPPLEMENTARY ANALYSES

Creates a histogram describing all the hospitalizations, divided by patients who have ever recieved a mental illness diagnosis, and patients who haven't
```{r}
no_mental=filter(encounter_diagnosis_patients,encounter_diagnosis_patients$EverDiagnosed==0)
yes_mental=filter(encounter_diagnosis_patients,encounter_diagnosis_patients$EverDiagnosed==1)
no_mental=data.frame(DurableKey=table(no_mental$PatientDurableKey))
yes_mental=data.frame(DurableKey=table(yes_mental$PatientDurableKey))

combined <- bind_rows(
  yes_mental %>% mutate(Group = "Mental illness"),
  no_mental %>% mutate(Group = "No mental illness")
)
ggplot(combined, aes(x = DurableKey.Freq, fill = Group)) +
  geom_histogram(aes(y = after_stat(density)), 
                 alpha = 0.5, position = "identity") +
  xlim(0, 175) +
  labs(title = "Frequency of visits by patients",
       x = "Frequency",
       y = "Proportion") 
```

Creates a histogram displaying the MyChart inactivation rate  by birth year. *No significant results.*

```{r}
# encounter_diagnosis_patients <- encounter_diagnosis_patients %>%
#   mutate(activated_flag = if_else(MyChartStatus == "Activated", 1, 0))
# 
# ggplot(encounter_diagnosis_patients, 
#        aes(x = PatientBirthYearBin, y = activated_flag)) +
#   stat_summary(fun = mean, geom = "col", fill = "cadetblue") +
#   labs(title = "MyChart inactivation rate by birth year",
#        x = "Birth Year",
#        y = "Proportion inactivated") +
#   theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
#   scale_x_continuous(breaks = seq(1920, 2025, 10)) + xlim(1900,2025)
```
Creates a line graph displaying repeat hospitalizations in patients diagnosed with mental disorders. *No significant results.*

```{r}
# encounter_diagnosis_patients <- encounter_diagnosis_patients %>%
#   mutate(Date=mdy(Date))
# patient_counts=as.data.frame(table(addiction_hospital$PatientDurableKey))
# freq_df <- patient_counts %>%
#   count(Freq)
# ggplot(freq_df, aes(x = Freq, y = n)) +
#   geom_line(color = "cadetblue") +
#   geom_point(color = "cadetblue") +
#   labs(title = "Repeat Hospitalizations Distribution in Patients with Diagnosed Mental Disorders",
#        x = "Number of Hospitalizations",
#        y = "Number of Patients")
```

Just played around with the dataset but didn't find much. *No significant results.*
```{r}
# addiction = filter(encounter_diagnosis_patients,
#                    GroupCode==   ("ICD-10-CM: F", 10:19))
# addiction_hospital <- addiction %>%
#   filter(AdmissionType == "Emergency") %>%
#   arrange(PatientDurableKey)
# addiction_key <- addiction_hospital %>%
#   count(PatientDurableKey, name = "Frequency")
# 
# addiction_hospital_final <- addiction %>%
#   inner_join(addiction_key, by = "PatientDurableKey")
```

Creates a graph displaying distribution of all races in dataset. *No use as of yet*

```{r}
# patient_stats <- encounter_diagnosis_patients %>%
#   distinct(PatientDurableKey, FirstRace, EverDiagnosed)
# race_prevalence <- patient_stats %>%
#   group_by(FirstRace) %>%
#   summarize(
#     Total_Patients = n(),
#     Diagnosed_Count = sum(EverDiagnosed),
#     Prevalence_Pct = mean(EverDiagnosed) * 100
#   ) %>%
#   arrange(desc(Prevalence_Pct))
# race_prevalence=race_prevalence%>%slice(-c(7,8,10,11))
# ggplot(race_prevalence, aes(x = reorder(FirstRace, Prevalence_Pct), y = Prevalence_Pct)) +
#   geom_col(fill = "orchid4") +
#   coord_flip() +
#   labs(title = "Prevalence of Mental Illness by Race",
#        x = "Race",
#        y = "Prevalence")

```

Follow up and mychart activity sample plots. *No use as of yet.*
```{r}
#FOR MENTALLY ILL PATIENTS
mental=filter(encounter_diagnosis_patients,EverDiagnosed==1)
analysis_df1 <- mental %>%
  mutate(
    Is_FollowUp = if_else(VisitType == "FOLLOW UP", 1, 0),
        Simple_Status = if_else(MyChartStatus == "Activated", "Active", "Not Active")
  ) %>%
  filter(!is.na(Simple_Status), !is.na(VisitType))

# 1. Create a contingency table
contingency_table1 <- table(analysis_df1$Simple_Status, analysis_df1$Is_FollowUp)
print(contingency_table1)

# 2. Run the Chi-Squared Test
chi_test1 <- chisq.test(contingency_table1)
print(chi_test1)

followup_comparison1 <- analysis_df1 %>%
  group_by(Simple_Status) %>%
  summarize(
    Total_Encounters = n(),
    FollowUp_Count = sum(Is_FollowUp),
    FollowUp_Rate = (FollowUp_Count / Total_Encounters) * 100
  )

print(followup_comparison1)

ggplot(followup_comparison1, aes(x = Simple_Status, y = FollowUp_Rate, fill = Simple_Status)) +
  geom_col(width = 0.6) +
  geom_text(aes(label = paste0(round(FollowUp_Rate, 1), "%")), vjust = -0.5) +
  labs(
    y = "Follow-up Rate (%)",
    x = "MyChart Status"
  ) +
  theme_minimal() +
  guides(fill = "none")


#FOR NON MENTALLY ILL PATIENTS
nonmental=filter(encounter_diagnosis_patients,EverDiagnosed==0)
analysis_df2 <- nonmental %>%
  mutate(
    Is_FollowUp = if_else(VisitType == "FOLLOW UP", 1, 0),
    
    # Simplify MyChart status into two main groups
    Simple_Status = if_else(MyChartStatus == "Activated", "Active", "Not Active")
  ) %>%
  # Remove NAs to keep the statistical test clean
  filter(!is.na(Simple_Status), !is.na(VisitType))
# 1. Create a contingency table
contingency_table2 <- table(analysis_df2$Simple_Status, analysis_df2$Is_FollowUp)
print(contingency_table2)

# 2. Run the Chi-Squared Test
chi_test2 <- chisq.test(contingency_table2)
print(chi_test2)

followup_comparison2 <- analysis_df2 %>%
  group_by(Simple_Status) %>%
  summarize(
    Total_Encounters = n(),
    FollowUp_Count = sum(Is_FollowUp),
    FollowUp_Rate = (FollowUp_Count / Total_Encounters) * 100
  )

#FOR BOTH BUT WITH NO DIVISION. NOTE; NO USE AS OF YET
# print(followup_comparison2)
# analysis_df1 <- encounter_diagnosis_patients %>%
#   mutate(
#     Is_FollowUp = if_else(VisitType == "FOLLOW UP", 1, 0),
#         Simple_Status = if_else(MyChartStatus == "Activated", "Active", "Not Active")
#   ) %>%
#   filter(!is.na(Simple_Status), !is.na(VisitType))
# 
# # 1. Create a contingency table
# contingency_table <- table(analysis_df$Simple_Status, analysis_df$Is_FollowUp)
# print(contingency_table)
# 
# # 2. Run the Chi-Squared Test
# chi_test <- chisq.test(contingency_table)
# print(chi_test)
# 
# followup_comparison <- analysis_df %>%
#   group_by(Simple_Status) %>%
#   summarize(
#     Total_Encounters = n(),
#     FollowUp_Count = sum(Is_FollowUp),
#     FollowUp_Rate = (FollowUp_Count / Total_Encounters) * 100
#   )
# 
# print(followup_comparison)
# 
# ggplot(followup_comparison, aes(x = Simple_Status, y = FollowUp_Rate, fill = Simple_Status)) +
#   geom_col(width = 0.6) +
#   geom_text(aes(label = paste0(round(FollowUp_Rate, 1), "%")), vjust = -0.5) +
#   labs(
#     title = "Follow-up Rate: Active vs. Not Active MyChart Users",
#     y = "Follow-up Rate (%)",
#     x = "MyChart Status"
#   ) +
#   guides(fill = "none") 
```

Looks for relationship between relationship status and myChart activation status. Strong relationship found but not used in analysis. *No use as of yet.*
```{r}
marital_analysis <- encounter_diagnosis_patients %>%
  mutate(MaritalStatus_Clean = case_when(
    MaritalStatus == "Single" ~ "Single",
    MaritalStatus == "Married" ~ "Married",
    MaritalStatus == "Widowed" ~ "Widowed",
    MaritalStatus %in% c("Divorced", "Legally Separated") ~ "Divorced",
    MaritalStatus %in% c("Significant Other", "Life Partner") ~ "Significant Other",
    TRUE ~ NA_character_ 
  )) %>%
  filter(!is.na(MaritalStatus_Clean))

# 2. Calculate Activation Rates
marital_stats <- marital_analysis %>%
  group_by(MaritalStatus_Clean) %>%
  summarize(
    Total_Patients = n_distinct(PatientDurableKey),
    Activated_Count = sum(activated_flag == 1, na.rm = TRUE),
    Activation_Rate = (Activated_Count / Total_Patients)
  ) %>%
  arrange(desc(Activation_Rate))

print(marital_stats)

marital_table <- table(marital_analysis$MaritalStatus_Clean, marital_analysis$activated_flag)
marital_proportions <- prop.table(marital_table, margin = 1)
print(marital_table)
marital_chisq <- chisq.test(marital_table)
print(marital_chisq)
ggplot(marital_stats, aes(x = reorder(MaritalStatus_Clean, Activation_Rate), y = Activation_Rate, fill = MaritalStatus_Clean)) +
  geom_col() +
  coord_flip() +
  geom_text(aes(label = paste0(round(Activation_Rate, 1), "%")), hjust = -0.1) +
  labs(
    title = "MyChart Activation by Marital Status",
    subtitle = "Testing the Social Support Hypothesis",
    x = "Marital Category",
    y = "Activation Rate (%)"
  )
```

Creates a plot showing top diagnosed illnesses and myChart inactivity. Found strong relationship between mental illness (6/10 top cases) and inactivity.
```{r}
disease_inactivation_analysis <- encounter_diagnosis_patients %>%
  filter(!is.na(DiagnosisName)) %>%
  group_by(DiagnosisName) %>%
  summarize(
    Total_Patients = n_distinct(PatientDurableKey),
    Inactivated_Patients = n_distinct(PatientDurableKey[activated_flag==0]),
    Inactivation_Rate = (Inactivated_Patients / Total_Patients) * 100
  ) %>%
  filter(Total_Patients > 500) %>% 
  arrange(desc(Inactivation_Rate)) %>%
  slice_head(n = 10) 
disease_inactivation_analysis <- disease_inactivation_analysis %>%
  mutate(DiagnosisName=c("Disruptive mood dysregulation disorder 
","Major depressive disorder with psychotic symptoms
","Examination for participation in sport
","Suicidal ideations
","Respiratory tuberculosis screening
","Major depressive disorder without psychotic features
","Short stature (children)
","Encounter for pre-employment examination
","Altered mental status, unspecified
","Major depressive disorder (single episode)
"))
plot <- ggplot(disease_inactivation_analysis, aes(x = reorder(DiagnosisName, Inactivation_Rate), y = Inactivation_Rate)) +
  geom_col(fill = "#d95f02", width = 0.7) + 
  coord_flip() +
    geom_text(aes(label = paste0(round(Inactivation_Rate, 1), "%")), 
            hjust = -0.2, 
            size = 3) +
    scale_x_discrete(labels = function(x) str_wrap(x, width = 25)) + 
    scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Top 10 Diagnoses by MyChart Inactivation",
    x = NULL, # Remove "Diagnosis" label to save space
    y = "Inactivation Rate (%)"
  ) +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 8, lineheight = 0.8), # Small text with tight line spacing
    axis.text.x = element_text(size = 8),
    plot.title = element_text(size = 10, face = "bold", hjust = 0.5),
    panel.grid.major.y = element_blank(), # Remove horizontal lines for a cleaner look
    plot.margin = margin(10, 20, 10, 10) # Add extra space on the right for the % labels
  )
ggsave("readable_inactivation_plot.png", plot, width = 6, height = 8, dpi = 300)
```