---
title: "datafest analysis"
output: html_document
date: "2026-03-28"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(scipen=999)

# Load required packages
library(tidyverse)

# Read datasets (using relative paths for GitHub portability)
encounters=read.csv("encounters.csv")
diagnosis=read.csv("diagnosis.csv")
patients=read.csv("patients.csv")
social_determinants=read.csv("social_determinants.csv")
departments=read.csv("departments.csv")
tigercensuscodes=read.csv("tigercensuscodes.csv")
```

## Join datasets to create a big one with encounters, diagnosis and patients. Also creates the variable EverDiagnosed, which is 1 if a patient was, throughout their journey, diagnosed with a mental illness or disorder.

```{r}
#unique primary diagnosis key
diagnosis_unique = diagnosis %>%
  distinct(DiagnosisKey, .keep_all = TRUE)

encounter_diagnosis = left_join(
  encounters,
  diagnosis_unique,
  by = c("PrimaryDiagnosisKey" = "DiagnosisKey")
)
encounter_diagnosis_patients=left_join(encounter_diagnosis,patients,by=c("PatientDurableKey"="DurableKey"))

encounter_diagnosis_patients <- encounter_diagnosis_patients %>% 
  mutate(MentalDisorder = ifelse(str_starts(DiagnosisValue, "F") & !str_starts(DiagnosisValue, "F7") | str_starts(DiagnosisValue, "R45"),1,0))

encounter_diagnosis_patients <- encounter_diagnosis_patients %>%
  group_by(PatientDurableKey) %>%
  mutate(
    EverDiagnosed = if_else(any(MentalDisorder == 1, na.rm = TRUE), 1, 0)
  ) %>%
  ungroup()

encounter_diagnosis_patients <- encounter_diagnosis_patients %>%
  group_by(PatientDurableKey) %>%
  summarize(
    # Returns TRUE if any encounter for this patient was a 1
    EverDiagnosed = any(MentalDisorder == 1, na.rm = TRUE)
  ) %>%
  mutate(EverDiagnosed = as.numeric(EverDiagnosed))

```

Creates a histogram describing hospitalizations, grouped by patients who have ever received a mental illness diagnosis versus those who haven't.

```{r}
no_mental=filter(encounter_diagnosis_patients,encounter_diagnosis_patients$EverDiagnosed==0)
yes_mental=filter(encounter_diagnosis_patients,encounter_diagnosis_patients$EverDiagnosed==1)
no_mental=data.frame(DurableKey=table(no_mental$PatientDurableKey))
yes_mental=data.frame(DurableKey=table(yes_mental$PatientDurableKey))

combined <- bind_rows(
  yes_mental %>% mutate(Group = "Mental illness"),
  no_mental %>% mutate(Group = "No mental illness")
)
ggplot(combined, aes(x = DurableKey.Freq, fill = Group)) +
  geom_histogram(aes(y = after_stat(density)), 
                 alpha = 0.5, position = "identity") +
  xlim(0, 175) +
  labs(title = "Frequency of visits by patients",
       x = "Frequency",
       y = "Proportion") 
```

Creates a histogram displaying the MyChart inactivation rate  by birth year. *No significant results.*

```{r}
# encounter_diagnosis_patients <- encounter_diagnosis_patients %>%
#   mutate(activated_flag = if_else(MyChartStatus == "Activated", 1, 0))
# 
# ggplot(encounter_diagnosis_patients, 
#        aes(x = PatientBirthYearBin, y = activated_flag)) +
#   stat_summary(fun = mean, geom = "col", fill = "cadetblue") +
#   labs(title = "MyChart inactivation rate by birth year",
#        x = "Birth Year",
#        y = "Proportion inactivated") +
#   theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
#   scale_x_continuous(breaks = seq(1920, 2025, 10)) + xlim(1900,2025)
```
Creates a line graph displaying repeat hospitalizations in patients diagnosed with mental disorders. *No significant results.*

```{r}
# encounter_diagnosis_patients <- encounter_diagnosis_patients %>%
#   mutate(Date=mdy(Date))
# patient_counts=as.data.frame(table(addiction_hospital$PatientDurableKey))
# freq_df <- patient_counts %>%
#   count(Freq)
# ggplot(freq_df, aes(x = Freq, y = n)) +
#   geom_line(color = "cadetblue") +
#   geom_point(color = "cadetblue") +
#   labs(title = "Repeat Hospitalizations Distribution in Patients with Diagnosed Mental Disorders",
#        x = "Number of Hospitalizations",
#        y = "Number of Patients")
```

Creates a graph displaying distribution of all races in dataset. *No use as of yet*

```{r}
# patient_stats <- encounter_diagnosis_patients %>%
#   distinct(PatientDurableKey, FirstRace, EverDiagnosed)
# race_prevalence <- patient_stats %>%
#   group_by(FirstRace) %>%
#   summarize(
#     Total_Patients = n(),
#     Diagnosed_Count = sum(EverDiagnosed),
#     Prevalence_Pct = mean(EverDiagnosed) * 100
#   ) %>%
#   arrange(desc(Prevalence_Pct))
# race_prevalence=race_prevalence%>%slice(-c(7,8,10,11))
# ggplot(race_prevalence, aes(x = reorder(FirstRace, Prevalence_Pct), y = Prevalence_Pct)) +
#   geom_col(fill = "orchid4") +
#   coord_flip() +
#   labs(title = "Prevalence of Mental Illness by Race",
#        x = "Race",
#        y = "Prevalence")

```


